---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# ggGWAS 🚧


<!-- badges: start -->
[![Travis build status](https://travis-ci.org/sinarueeger/ggGWAS.svg?branch=master)](https://travis-ci.org/sinarueeger/ggGWAS)
[![Aodecov test coverage](https://codecov.io/gh/sinarueeger/ggGWAS/branch/master/graph/badge.svg)](https://codecov.io/gh/sinarueeger/ggGWAS?branch=master)
<!-- badges: end -->


An R-Package (*work-in-progress*) that contains ggplot2-extensions of data visualisations used with GWAS data. 

Mainly, these are Q-Q plot and Manhattan plot that both use P-values from GWASs as input. 

An inspiration for ggGWAS has been the R-package [qqman](http://www.gettinggeneticsdone.com/2014/05/qqman-r-package-for-qq-and-manhattan-plots-for-gwas-results.html), except that ggGWAS aims to have the look and functionality of `ggplot2`.



## Installation

You can install the development version from [GitHub](https://github.com/) with:


```r
# install.packages("remotes")
remotes::install_github("sinarueeger/ggGWAS")
```

Install the `dev`b ranch:
```r
remotes::install_github("sinarueeger/ggGWAS", ref = "dev")
```


Install including vignettes: 

```r
remotes::install_github("sinarueeger/ggGWAS", build = TRUE, build_opts = c("--no-resave-data", "--no-manual"))
vignette("gggwas-reasoning")
vignette("gggwas-internals")
```


## Basic usage

```{r data}
library(ggGWAS)
library(ggplot2)

theme_set(theme_bw())
## Generate some random data

df <-
  data.frame(
    POS = rep(1:250, 4),
    CHR = 1:4,
    P = runif(1000),
    GWAS = sample(c("a", "b"), 1000, replace = TRUE)
  )

```



### Manhattan plot

```{r mhtplot}
ggplot(data = df) + stat_gwas_manhattan(aes(pos = POS, y = -log10(P), chr = CHR))

?stat_gwas_manhattan ## for more examples
```

### Q-Q plot

Lightweight Q-Q plot (with hex tiles)

```{r qqplot-hex}

ggplot(data = df) + stat_gwas_qq_hex(aes(y = P))

?stat_gwas_qq_hex ## for more examples
```

Conventional Q-Q plot
```{r qqplot}
ggplot(data = df) + stat_gwas_qq(aes(y = P))

?stat_gwas_qq ## for more examples
```


## Contributing
 lease note that the 'ggGWAS' project is released with a
[Contributor Code of Conduct](CODE_OF_CONDUCT.md).
By contributing to this project, you agree to abide by its terms.
