---
title: ggGWAS Internals
author: Sina Rueger
date: "`r Sys.Date()`"
show_toc: true
slug: gggwas-internals
output: 
  rmarkdown::html_vignette:
      toc: yes
vignette: >
  %\VignetteIndexEntry{ggGWAS Internals}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(ggplot2)
library(ggGWAS)
library(GWAS.utils)
```
This vignette contains some technical aspects about developing the `ggGWAS` package. 


## New scale `-log10`

There are different scales available in ggplot2.

- `scale_x_log10` and `scale_y_log10`
- `scale_x_continuous(trans = "log10")` and `scale_y_continuous(trans = "log10")`
- `coord_trans(x = "log10", y = "log10")`

```{r}
library(ggplot2)
df <- data.frame(y = runif(1000))

qp <- ggplot(df, aes(sample = y)) +
  stat_qq(distribution = stats::qunif)

qp + scale_x_log10() + scale_y_log10()
qp + scale_x_continuous(trans = "log10") + scale_y_continuous(trans = "log10")
qp + coord_trans(x = "log10", y = "log10")

```

For our Q-Q plot we want to create a new transformation called `-log10`. 


```{r}

ggplot(giant, aes(sample = P)) +
   stat_qq(distribution = stats::qunif) +
     scale_y_continuous(trans = ggGWAS::mlog10_trans()) +
   scale_x_continuous(trans = ggGWAS::mlog10_trans()) +
   geom_abline(intercept = 0, slope = 1)

```


## Speed

One aspect of `ggGWAS` is improving speed. Even though `ggplot2` itself is not known for its speed, we can apply a few tricks to beat the base plot speed. 

1. Using rastering (`ggrastr:::GeomPointRast`)
2. Using hexagons (`ggplot::geom_hex`)
3. Subsetting data
4. Avoiding passing on the full datasets (???)

To compare, we use the time to plot in base graphics (`qqman` package):


```{r}
n.sample <- 1e3
set.seed(3)
df <- data.frame(P = runif(n.sample), GWAS = sample(c("a","b"), n.sample, replace = TRUE))
system.time(qqman::qq(df$P))
```


### 1. Raster

The R-package `ggrastr` does ???. 

It creates very light output, but is not necessarily fast. 

Here is a comparison of the actual pdf file size.

```{r eval=FALSE, include=FALSE}

## >>>>>>> just check if ggrastr is installed

## raster
gg_vec <- ggplot(df, aes(x = P, y = P)) + geom_point()
gg_ras <- gg_vec + ggrastr::geom_point_rast(size = 0.5)

PrintFileSize <- function(gg, name) {
  invisible(ggsave("tmp.pdf", gg, width = 4, height = 4))
  cat(name, ": ", file.info("tmp.pdf")$size / 1024, " Kb.\n", sep = "")
  unlink("tmp.pdf")
}

PrintFileSize(gg_ras, "Raster")
PrintFileSize(gg_vec, "Vector")

```

We can use the rastering functionality as geom in our function. 

```{r eval=FALSE, include=FALSE}

qp_points <- ggplot(df, aes(observed = P)) +
  stat_gwas_qq() +
  geom_abline(intercept = 0, slope = 1)
system.time(print(qp_points))

qp_raster <- ggplot(df, aes(observed = P)) +
  stat_gwas_qq(geom = ggrastr:::GeomPointRast) +
  geom_abline(intercept = 0, slope = 1)
system.time(print(qp_raster))

PrintFileSize(qp_points, "Vector")
PrintFileSize(qp_raster, "Raster")


```

### 2. Hexagons

```{r}

qp_hex <- ggplot(df, aes(y = P)) +
  stat_gwas_qq_hex() +
  geom_abline(intercept = 0, slope = 1)
system.time(print(qp_hex))
```

Problem is, that the original dataset is still passed on. This is not a problem in terms of speed, but can be for memeory. 

```{r}
#library(pryr)
#object_size(gg)

```

The default `hex.function` `hexBinSummarise_custom` does not have the usual properties. Will have the hexagon shape, but not any count. Tweaked `hexbinheight`.

`hexBinSummarise` will have the usual properties.
```{r}
#gg <- ggplot(data = df) + stat_gwas_qq_hex(aes(y = P), fill = "black", hex.function = hexBinSummarise) # 418x13 table in gb$data

```


### 3. Avoiding passing on the full datasets (???)


```{r}

ggplot_build(qp_hex) ## data structure

ggplot_gtable(ggplot_build(qp_hex)) ## grid object

```


### 4. Subsetting data

Something very simple is to subset the data, since we are usually not interested in the very high P-values (although, we are at times, when it comes to model sanity checks).


```{r}
qp_subset <- ggplot(df, aes(observed = P)) +
  stat_gwas_qq(observed.thresh = 0.05) +
  geom_abline(intercept = 0, slope = 1)
system.time(print(qp_subset))


```

## New color palette

See also: https://github.com/mkanai/ggman/blob/master/R/scale-colour-dichromatic.R

## New theme

See also: https://github.com/mkanai/ggman/blob/master/R/theme-publication.R



